---
title: Investigating the impact of the COVID-19 pandemic on the gender productivity
  gap in academia
author: "Kiran Gok Lune Lee"
date: "`r Sys.Date()`"
geometry: margin=2cm 
output:
  pdf_document:
    latex_engine: xelatex
    fig_width: 8
    fig_height: 6
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    fig_width: 12
    fig_height: 10

---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

```


## Install and load packages

```{r eval = FALSE, echo = TRUE}
## install packages like so
install.packages("pacman")
rm(list = ls())
devtools::install_github("daniel1noble/orchaRd", force = TRUE)
pacman::p_load(devtools, tidyverse, metafor, patchwork, R.rsp, orchaRd, emmeans, ape, phytools, flextable)



```

```{r eval = TRUE, echo = FALSE}
library(metafor)
library(readxl)
library(ggplot2)
library(formatR)
library(devtools)
library(tidyverse)
library(patchwork)
library(orchaRd)
library(gridExtra)

```


## Load data
```{r eval = TRUE, echo = FALSE}
all_data<-read_xlsx("all_data.xlsx")
colnames(all_data) <- make.names(colnames(all_data), unique = TRUE)


### set effect sizes
all_data$Effect.size.used.in.MA<-as.numeric(all_data$Effect.size.used.in.MA)
all_data$Variance.used<-as.numeric(all_data$Variance.used)
all_data$vi <- all_data$Variance.used
all_data$yi <- all_data$Effect.size.used.in.MA

### data used for writing methods
table(all_data$Effect.size.kiran.use..from.Campbell.collaboration.)
table(all_data$Self.reported.or.measured)
table(all_data$Broad.research.field)
table(all_data$Broad.productivity.measure)
table(all_data$Broad.specific.productivity)
table(all_data$Specific.productivity.measure)

all_data<-arrange(all_data, Effect.size.used.in.MA) # Order the data by effect size ID for plots
all_data$Effect.size.used.in.MA<-as.factor(all_data$Effect.size.used.in.MA)

### subset effect sizes using submission/publication output only
measured_data<-subset(all_data, all_data$Self.reported.or.measured == "Measured")
survey_data<-subset(all_data, all_data$Self.reported.or.measured == "Self-reported")

#Article ID refers to the article ID number the effect size comes from
all_data$ID.article<-as.factor(all_data$Article.ID)
#ID refers to the unique effect size ID number
all_data$ID.observation<-as.factor(all_data$ID)

#Average time frame before pandemic and during pandemic for article studies
measured_data$Timeframe.pre.pandemic<-as.numeric(measured_data$Timeframe.pre.pandemic)
mean(measured_data$Timeframe.pre.pandemic, na.rm= TRUE)
sd(measured_data$Timeframe.pre.pandemic, na.rm = TRUE)
range(measured_data$Timeframe.pre.pandemic, na.rm = TRUE)

measured_data$Timeframe.during.pandemic<-as.numeric(measured_data$Timeframe.during.pandemic)
mean(measured_data$Timeframe.during.pandemic, na.rm = TRUE)
sd(measured_data$Timeframe.during.pandemic, na.rm = TRUE)
range(measured_data$Timeframe.during.pandemic, na.rm = TRUE)

#Average end date for survey studies
survey_data$End.date.of.survey<-as.numeric(survey_data$End.date.of.survey)
as.Date(mean(survey_data$End.date.of.survey, na.rm =TRUE), origin = "1899-12-30")
sd(survey_data$End.date.of.survey, na.rm =TRUE)
as.Date(range(survey_data$End.date.of.survey, na.rm =TRUE), origin = "1899-12-30")
as.Date(max(survey_data$End.date.of.survey, na.rm =TRUE), origin = "1899-12-30")-as.Date(min(survey_data$End.date.of.survey, na.rm =TRUE), origin = "1899-12-30")

#Geographic regions
table(all_data$Broad.region)
table(measured_data$Broad.region)
table(survey_data$Broad.region)
```


```{r eval = TRUE, echo = FALSE}

#Custom settings for orchaRd plots borrowed from https://github.com/p-pottier/Dev_plasticity_thermal_tolerance/blob/main/Data_analysis/R/Statistical_analyses.Rmd
my.orchard<- function (object, mod = "1", group, data, xlab, N = "none", 
    alpha = 0.5, angle = 0, cb = FALSE, k = TRUE, g = TRUE, 
    trunk.size = 7, branch.size = 2, twig.size = 0.8, whisker, transfm = c("none", # increased point size, branch size, and added a whisker argument
        "tanh"), condition.lab = "Condition", legend.pos = "bottom.right", k.pos = c("right", 
        "left")) 
{
    transfm <- match.arg(transfm)
    if (any(class(object) %in% c("rma.mv", "rma"))) {
        if (mod != "1") {
            results <- orchaRd::mod_results(object, mod, group, 
                data)
        }
        else {
            results <- orchaRd::mod_results(object, mod = "1", 
                group, data)
        }
    }
    if (any(class(object) %in% c("orchard"))) {
        results <- object
    }
    mod_table <- results$mod_table
    data <- results$data
    data$moderator <- factor(data$moderator, levels = mod_table$name, 
        labels = mod_table$name)
    data$scale <- (1/sqrt(data[, "vi"]))
    legend <- "Precision (1/SE)"
    if (any(N != "none")) {
        data$scale <- N
        legend <- "Sample Size (N)"
    }
    if (transfm == "tanh") {
        cols <- sapply(mod_table, is.numeric)
        mod_table[, cols] <- Zr_to_r(mod_table[, cols])
        data$yi <- Zr_to_r(data$yi)
        label <- xlab
    }
    else {
        label <- xlab
    }
    mod_table$K <- as.vector(by(data, data[, "moderator"], 
        function(x) length(x[, "yi"])))
    mod_table$g <- as.vector(num_studies(data, moderator, stdy)[, 
        2])
    group_no <- length(unique(mod_table[, "name"]))
    cbpl <- c("#88CCEE", "#CC6677", "#DDCC77", 
        "#117733", "#332288", "#AA4499", "#44AA99", 
        "#999933", "#882255", "#661100", "#6699CC", 
        "#888888", "#E69F00", "#56B4E9", "#009E73", 
        "#F0E442", "#0072B2", "#D55E00", "#CC79A7", 
        "#999999")
    if (names(mod_table)[2] == "condition") {
        condition_no <- length(unique(mod_table[, "condition"]))
        plot <- ggplot2::ggplot() + ggbeeswarm::geom_quasirandom(data = data, 
            ggplot2::aes(y = yi, x = moderator, size = scale, 
                colour = moderator), alpha = alpha) + ggplot2::geom_hline(yintercept = 0, 
            linetype = 2, colour = "black", alpha = alpha) + 
            ggplot2::geom_linerange(data = mod_table, ggplot2::aes(x = name, 
                ymin = lowerCL, ymax = upperCL), size = branch.size, 
                position = ggplot2::position_dodge2(width = 0.3)) + 
            ggplot2::geom_pointrange(data = mod_table, ggplot2::aes(y = estimate, 
                x = name, ymin = lowerPR, ymax = upperPR, shape = as.factor(condition), 
                fill = name), size = twig.size, position = ggplot2::position_dodge2(width = 0.3), 
                fatten = trunk.size) + ggplot2::scale_shape_manual(values = 20 + 
            (1:condition_no)) + ggplot2::coord_flip() + ggplot2::theme_bw() + 
            ggplot2::guides(fill = "none", colour = "none") + 
            ggplot2::theme(legend.position = c(0, 1), legend.justification = c(0, 
                1)) + ggplot2::theme(legend.title = ggplot2::element_text(size = 9)) + 
            ggplot2::theme(legend.direction = "horizontal") + 
            ggplot2::theme(legend.background = ggplot2::element_blank()) + 
            ggplot2::labs(y = label, x = "", size = legend) + 
            ggplot2::labs(shape = condition.lab) + ggplot2::theme(axis.text.y = ggplot2::element_text(size = 10, 
            colour = "black", hjust = 0.5, angle = angle))
    }
    else {
        plot <- ggplot2::ggplot() + ggbeeswarm::geom_quasirandom(data = data, 
            ggplot2::aes(y = yi, x = moderator, size = scale, 
                fill = moderator), alpha = alpha, width=0.4, pch=21, stroke=1.1, col="black") + # Change point shape (21, with black borders)
            ggplot2::geom_hline(yintercept = 0, 
            linetype = 2, colour = "black", alpha = 0.3, lwd=1.3) + # Change thickness 0 line
            ggplot2::geom_errorbar(data = mod_table, ggplot2::aes(x = name, 
                ymin = lowerCL, ymax = upperCL), size = branch.size, width= whisker) + # Added variable whisker size
            ggplot2::geom_pointrange(data = mod_table, ggplot2::aes(y = estimate, 
                x = name, ymin = lowerPR, ymax = upperPR, fill = name), 
                size = twig.size, fatten = trunk.size, shape = 23, stroke=2.2) + # Change point shape
            scale_size_continuous(range = c(1, 14))+ # change point scaling
            ggplot2::coord_flip() + 
            ggplot2::theme_bw() + 
            ggplot2::guides(fill = "none", colour = "none") + 
            ggplot2::theme(text=element_text(size=26, colour="black"))+ # Change font size
            ggplot2::theme(legend.title = ggplot2::element_text(size = 16)) + # Increased font legend title
            ggplot2::theme(legend.text = ggplot2::element_text(size = 14)) +
            ggplot2::theme(legend.direction = "horizontal") + 
            ggplot2::theme(legend.background = ggplot2::element_blank()) + 
            ggplot2::labs(y = label, x = "", size = legend) + 
            ggplot2::theme(axis.text.y = ggplot2::element_text(size = 20, 
                colour = "black", hjust = 0.5, angle = angle)) + # Increased size title axis label
            ggplot2::theme(axis.text.x = ggplot2::element_text(size = 20)) + # Increase size axis ticks
            ggplot2::theme(panel.border = element_rect(colour = "black", fill=NA, size=1.3))
            
    }
    if (legend.pos == "bottom.right") {
        plot <- plot + ggplot2::theme(legend.position = c(1, 
            0), legend.justification = c(1, 0))
    }
    else if (legend.pos == "bottom.left") {
        plot <- plot + ggplot2::theme(legend.position = c(0, 
            0), legend.justification = c(0, 0))
    }
    else if (legend.pos == "top.right") {
        plot <- plot + ggplot2::theme(legend.position = c(1, 
            1), legend.justification = c(1, 1))
    }
    else if (legend.pos == "top.left") {
        plot <- plot + ggplot2::theme(legend.position = c(0, 
            1), legend.justification = c(0, 1))
    }
    else if (legend.pos == "top.out") {
        plot <- plot + ggplot2::theme(legend.position = "top")
    }
    else if (legend.pos == "bottom.out") {
        plot <- plot + ggplot2::theme(legend.position = "bottom")
    }
    if (cb == TRUE) {
        plot <- plot + ggplot2::scale_fill_manual(values = cbpl) + 
            ggplot2::scale_colour_manual(values = cbpl)
    }
    if (k == TRUE && g == FALSE && k.pos == "right") {
        plot <- plot + ggplot2::annotate("text", y = (max(data$yi) + 
            (max(data$yi) * 0.1)), x = (seq(1, group_no, 1) + 
            0.3), label = paste("italic(k)==", mod_table$K[1:group_no]), # Size changed to 5.5
            parse = TRUE, hjust = "right", size = 6.5)
    }
    else if (k == TRUE && g == FALSE && k.pos == "left") {
        plot <- plot + ggplot2::annotate("text", y = (min(data$yi) + 
            (min(data$yi) * 0.1)), x = (seq(1, group_no, 1) + 
            0.3), label = paste("italic(k)==", mod_table$K[1:group_no]), 
            parse = TRUE, hjust = "left", size = 6.5) # Size changed to 5.5
    }
    else if (k == TRUE && g == TRUE && k.pos == "right") {
        plot <- plot + ggplot2::annotate("text", y = (max(data$yi) + 
            (max(data$yi) * 0.1)), x = (seq(1, group_no, 1) + 
            0.3), label = paste("italic(k)==", mod_table$K[1:group_no], 
            " (", mod_table$g[1:group_no], ")"), 
            parse = TRUE, hjust = "right", size = 6.5) # Size changed to 5.5
    }
    else if (k == TRUE && g == TRUE && k.pos == "left") {
        plot <- plot + ggplot2::annotate("text", y = (min(data$yi) + 
            (min(data$yi) * 0.1)), x = (seq(1, group_no, 1) + 
            0.3), label = paste("italic(k)==", mod_table$K[1:group_no], 
            " (", mod_table$g[1:group_no], ")"), 
            parse = TRUE, hjust = "left", size = 6.5) # Size changed to 5.5
    }
    return(plot)
}

```

\newpage

## Hypothesis 1: During pandemic conditions the gender gap in academic productivity has increased.

Prediction 1a: The pandemic has increased the gender gap in productivity (as indicated by an overall negative effect size).

Looks like an overall negative effect: during pandemic conditions the gender gap in academic productivity has increased. 


```{r eval = TRUE, echo = TRUE}

#Hierarchical mixed effect meta-analysis with all papers. Studies with multiple effect sizes are controlled for.
m<-rma.mv(yi, vi, random=~1|ID.article/ID.observation, data=all_data)
m
summary(m)

#Forest plot produced for visual representation.
png(file='forest.png')
forest (m,  slab=all_data$Author, xlim=c(-2,2), ylim=c(-1, 140), digits= 2, xlab="Raw proportion", mlab="Overall effect (46)", cex=0.4)
text(-1,122, "Author(s) and Year", pos=2, font=2, cex=0.8)
text(2,122, "Effect size [95% CI]", pos=2, font=2, cex=0.8)
while (!is.null(dev.list()))  dev.off()

```

### Figure 1

```{r, fig.width=12, fig.height=12}

my.orchard(m, mod="1", alpha=0.5,data = all_data, whisker=0.025, group = "Article.ID", xlab = "Effect size")+
     annotate("text", size=6, y=0.66, x=1.2, label=paste("-0.071 [-0.099, -0.043]"))+
  scale_fill_manual(values=c("#DC267F"))

ggsave("figure1.png", width=10, height=10, dpi=300)

```

Prediction 1b: Though studies can measure the type of research productivity as by individual surveys, numbers of submissions and numbers of publications, this does not influence the gender gap increase observed during the pandemic.

```{r eval = TRUE, echo = TRUE}
all_data$Broad.productivity.measure[all_data$Broad.productivity.measure=="Other"]<-"Survey"

tapply(all_data$ID.observation, all_data$Broad.productivity.measure, length)

all_data$Broad.productivity.measure.reordered<-factor(all_data$Broad.productivity.measure, c("Submissions", "Publications", "Survey"))

m.area <- rma.mv(yi, vi, mods= ~ Broad.productivity.measure.reordered, random=~1|ID.article/ID.observation, data=all_data)
m.area
summary(m.area)

m.area1 <- rma.mv(yi, vi, mods= ~ Broad.productivity.measure.reordered-1, random=~1|ID.article/ID.observation, data=all_data)
m.area1

```

### Figure 2

```{r eval = TRUE, echo = TRUE, fig.width=6, fig.height=6}

p1 <- my.orchard(m.area1, mod="Broad.productivity.measure.reordered", group = "Article.ID", data=all_data, xlab = "Effect size", whisker=0.05,transfm = "none")+
     annotate("text", size=6, y=-0.9, x=3.3, label=paste("-0.192 [-0.272, -0.113]"))+
     annotate("text", size=6, y=-0.9, x=2.3, label=paste("-0.047 [-0.085, -0.008]"))+
     annotate("text", size=6, y=-0.9, x=1.3, label=paste("-0.053 [-0.087, -0.018]"))+
  scale_fill_manual(values=c("#785EF0","#DC267F","#FE6100"))

p1 

ggsave("figure2.png", width=10, height=10, dpi=300)
```

\newpage

## Hypothesis 2: During pandemic conditions the gender gap in academic productivity has increased differentially across research fields.

Prediction 2a: The pandemic has increased the gender gap in academic productivity more in fields that [already had a previously greater gender gap](https://royalsocietypublishing.org/doi/10.1098/rsos.181566) because these lacked gender-equitable support measures to prevent female academics experiencing research production setbacks.

```{r eval = TRUE, echo = TRUE, fig.width=6, fig.height=6}
all_data$Broad.research.field.reordered<-factor(all_data$Broad.research.field, c("Biological sciences","TEMCP", "Medicine", "Social sciences", "Multidisciplinary"))

levels(all_data$Broad.research.field.reordered) <- gsub(" ", "\n", levels(all_data$Broad.research.field.reordered))

research_field <- metafor::rma.mv(yi = yi, V = vi, mods = ~Broad.research.field.reordered, random = list(~1 |
    Article.ID, ~1 | ID), data = all_data)
summary(research_field)

research_field1 <- metafor::rma.mv(yi = yi, V = vi, mods = ~Broad.research.field.reordered - 1, random = list(~1 |
    Article.ID, ~1 | ID), data = all_data)

summary(research_field1)
```
Prediction 2a Measured studies: The pandemic has increased the gender gap in academic productivity more in fields that [already had a previously greater gender gap](https://royalsocietypublishing.org/doi/10.1098/rsos.181566) because these lacked gender-equitable support measures to prevent female academics experiencing research production setbacks.

```{r eval = TRUE, echo = TRUE, fig.width=6, fig.height=6}
measured_data$Broad.research.field.reordered<-factor(measured_data$Broad.research.field, c("TEMCP", "Biological sciences","Multidisciplinary", "Medicine", "Social sciences"))

levels(measured_data$Broad.research.field.reordered) <- gsub(" ", "\n", levels(measured_data$Broad.research.field.reordered))

research_fieldm <- metafor::rma.mv(yi = yi, V = vi, mods = ~Broad.research.field.reordered, random = list(~1 |
    Article.ID, ~1 | ID), data = measured_data)
summary(research_fieldm)

research_field1m <- metafor::rma.mv(yi = yi, V = vi, mods = ~Broad.research.field.reordered - 1, random = list(~1 |
    Article.ID, ~1 | ID), data = measured_data)

summary(research_field1m)
```

### Figure 3

```{r eval = TRUE, echo = TRUE, fig.width=6, fig.height=6}

my.orchard(research_field1, mod = "Broad.research.field.reordered", group = "Article.ID", data = all_data, whisker=0.08, xlab = "Effect size",
    alpha = 0.5, transfm = "tanh",  cb = FALSE)+
     annotate("text", size=6, y=0.5, x=5.13, label=paste("-0.109 [-0.173, -0.045]"))+
     annotate("text", size=6, y=0.5, x=4.13, label=paste("-0.096 [-0.155, -0.038]"))+
     annotate("text", size=6, y=0.5, x=3.13, label=paste("-0.083 [-0.113, -0.031]"))+
     annotate("text", size=6, y=0.5, x=2.13, label=paste("-0.017 [-0.069, 0.034]"))+
     annotate("text", size=6, y=0.5, x=1.13, label=paste("0.007 [-0.051, 0.064]"))+
  scale_fill_manual(values=c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000"))

ggsave("figure3.png", width=10, height=10, dpi=300)

```
### Figure 3 Measured data

```{r eval = TRUE, echo = TRUE, fig.width=6, fig.height=6}

my.orchard(research_field1m, mod = "Broad.research.field.reordered", group = "Article.ID", data = measured_data, whisker=0.08, xlab = "Effect size",
    alpha = 0.5, transfm = "tanh",  cb = FALSE)+
     annotate("text", size=6, y=-0.45, x=5.3, label=paste("-0.084 [-0.143, -0.024]"))+
     annotate("text", size=6, y=-0.45, x=4.3, label=paste("-0.066 [-0.102, -0.029]"))+
     annotate("text", size=6, y=-0.45, x=3.3, label=paste("-0.050 [-0.125, 0.024]"))+
     annotate("text", size=6, y=-0.45, x=2.3, label=paste("-0.003 [-0.057, 0.050]"))+
     annotate("text", size=6, y=-0.45, x=1.3, label=paste("-0.003 [-0.053, 0.048]"))+
  scale_fill_manual(values=c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000"))

ggsave("figure3.png", width=10, height=10, dpi=300)

```
Prediction 2b: We predict the pandemic has increased the gender gap more for research fields of a given authorship position that already had a previously greater gender gap because less support may have been available to women to balance the effects of the pandemic. 

```{r eval = TRUE, echo = TRUE, fig.width=8, fig.height=8}
all_data$nwomenprepandemic<-round(as.numeric(all_data$X..women.authors.pre.pandemic)*as.numeric(all_data$n.pre.pandemic), digits=0)
all_data$nmenprepandemic<-round(as.numeric(all_data$n.pre.pandemic)-as.numeric(all_data$nwomenprepandemic), digits=0)
all_data$nwomenduringpandemic<-round(as.numeric(all_data$X..women.authors.during.pandemic)*as.numeric(all_data$n.during.pandemic), digits=0)
all_data$nmenduringpandemic<-round(as.numeric(all_data$n.during.pandemic)-as.numeric(all_data$nwomenduringpandemic), digits=0)

previous_bias <- metafor::rma.mv(yi = yi, V = vi, mods = ~ cbind(nmenprepandemic/nwomenprepandemic), random = list(~1 |
    Article.ID, ~1 | ID), data = subset(all_data, all_data$Self.reported.or.measured=="Measured"))

#Model suggests that contrary to our prediction, the pandemic has increased the gender gap more for research fields of a given authorship position that were previously less biased.
summary(previous_bias)

measured_data_numbers<-measured_data
measured_data_numbers$X..women.authors.pre.pandemic<-as.numeric(measured_data_numbers$X..women.authors.pre.pandemic)
measured_data_numbers$X..women.authors.during.pandemic<-as.numeric(measured_data_numbers$X..women.authors.during.pandemic)
measured_data_numbers<-subset(measured_data_numbers, !is.na(measured_data_numbers$X..women.authors.pre.pandemic))
measured_data_numbers<-subset(measured_data_numbers, !is.na(measured_data_numbers$X..women.authors.during.pandemic))
socialsciences<-subset(measured_data_numbers, measured_data_numbers$Broad.research.field=="Social sciences")
medicine<-subset(measured_data_numbers, measured_data_numbers$Broad.research.field=="Medicine")
multi<-subset(measured_data_numbers, measured_data_numbers$Broad.research.field=="Multidisciplinary")
temcp<-subset(measured_data_numbers, measured_data_numbers$Broad.research.field=="TEMCP")
bio<-subset(measured_data_numbers, measured_data_numbers$Broad.research.field=="Biological sciences")

socialsciences$X..women.authors.pre.pandemic<-as.numeric(socialsciences$X..women.authors.pre.pandemic)
socialsciences$X..women.authors.during.pandemic<-as.numeric(socialsciences$X..women.authors.during.pandemic)
mean((socialsciences$X..women.authors.pre.pandemic))
mean((socialsciences$X..women.authors.during.pandemic))
mean((medicine$X..women.authors.pre.pandemic))
mean((medicine$X..women.authors.during.pandemic))
mean((multi$X..women.authors.pre.pandemic))
mean((multi$X..women.authors.during.pandemic))
mean((temcp$X..women.authors.pre.pandemic))
mean((temcp$X..women.authors.during.pandemic))
mean((bio$X..women.authors.pre.pandemic))
mean((bio$X..women.authors.during.pandemic))
mean((measured_data_numbers$X..women.authors.pre.pandemic))
mean((measured_data_numbers$X..women.authors.during.pandemic))

all_data$Broad.research.field.reordered.opposite<-factor(all_data$Broad.research.field, c( "Social sciences","Medicine","Multidisciplinary","Biological sciences","TEMCP" ))
all_data_long<-all_data
all_data_long$X..women.authors.before.pandemic<-all_data_long$X..women.authors.pre.pandemic
all_data_long$X..women.authors.before.pandemic<-as.numeric(all_data_long$X..women.authors.before.pandemic)
all_data_long$X..women.authors.during.pandemic<-as.numeric(all_data_long$X..women.authors.during.pandemic)

all_data_long <- pivot_longer(all_data_long, c(X..women.authors.before.pandemic, X..women.authors.during.pandemic), names_to = "Period")

#This plot shows the above model about how previous gender bias influence gender bias during the pandemic
ggplot(all_data_long, aes(x=factor(Period), y=as.numeric(value), color=factor(Broad.research.field),  group = ID)) + 
  geom_point(position = position_jitter(width = .0)) + 
  geom_smooth(method = 'lm', se = FALSE) + 
  labs(
    x = "x",
    color = "Broad.research.field"
  ) 


all_data_long$Period[all_data_long$Period == 'X..women.authors.before.pandemic']<-'Before pandemic'
all_data_long$Period[all_data_long$Period == 'X..women.authors.during.pandemic']<-'During pandemic'
all_data_long$'Broad research field'<-all_data_long$Broad.research.field.reordered

```

### Figure 4

```{r eval = TRUE, echo = TRUE, fig.width=8, fig.height=8}

ggplot(all_data_long, aes(x=factor(Period), y=as.numeric(value), color=factor(Broad.research.field.reordered.opposite),  group = Broad.research.field.reordered.opposite)) + 
  geom_point(position = position_jitter(width = .0)) + 
  geom_smooth(method = 'lm', se = FALSE) + 
  labs(
    color = "Research field"
  ) +
  labs( x = "", y = "Proportion female authors")+ theme_linedraw()+
  theme( panel.grid.major.x = element_blank() , panel.grid.major = element_line(size = 0.1, linetype = 'solid',
                                colour = "gray"), 
  panel.grid.minor = element_line(size = 0.1, linetype = 'solid',
                                colour = "gray"))+theme(axis.text=element_text(size=10))+scale_color_manual(values=c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000"))

ggsave("figure4.png", width=5, height=5, dpi=300)


```

\newpage

## Hypothesis 3: Pandemic conditions made it difficult for women to lead research, rather than support research.

Prediction 3: We predict the pandemic has increased the gender gap more in first and last, rather than middle authorship positions as female academics have been more limited in undertaking leading research roles, but not supportive research roles in pandemic conditions.

Note: Only 3 effect sizes from 2 studies that use middle authorship makes it difficult to test this.

```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}

measured_data$Specific.productivity.measure[measured_data$Specific.productivity.measure=="Sole authorship"]  <- "Last authorship"


measured_data$Specific.productivity.measure.reordered<-factor(measured_data$Specific.productivity.measure, c("Last authorship", "Middle authorship", "First authorship", "Corresponding authorship", "Any authorship"))

levels(measured_data$Specific.productivity.measure.reordered) <- gsub(" ", "\n", levels(measured_data$Specific.productivity.measure.reordered))

#Models
authorship_position <- metafor::rma.mv(yi = yi, V = vi, mods = ~Specific.productivity.measure.reordered, random = list(~1 | Article.ID, ~1 | ID), data = measured_data)

summary(authorship_position)

authorship_position1 <- metafor::rma.mv(yi = yi, V = vi, mods = ~Specific.productivity.measure.reordered - 1, random = list(~1 |
    Article.ID, ~1 | ID), data = measured_data)

summary(authorship_position1)
```

### Figure 5

```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}

my.orchard(authorship_position, mod = "Specific.productivity.measure.reordered", group = "Article.ID", data = measured_data, whisker = 0.08, xlab = "Effect size",
    alpha = 0.5)+
     annotate("text", size=6, y=-0.56, x=1.3, label=paste("-0.052 [-0.130, 0.026]"))+
     annotate("text", size=6, y=-0.56, x=2.3, label=paste("-0.046 [-0.178, 0.087]"))+
     annotate("text", size=6, y=-0.56, x=3.3, label=paste("-0.040 [-0.074, -0.006]"))+
     annotate("text", size=6, y=-0.56, x=4.3, label=paste("-0.041 [-0.122, 0.040]"))+
     annotate("text", size=6, y=-0.56, x=5.3, label=paste("-0.052 [-0.130, -0.026]"))+
  scale_fill_manual(values=c("#648FFF","#785EF0","#DC267F","#FE6100","#FFB000"))

ggsave("figure5.png", width=10, height=10, dpi=300)
```

\newpage

## Heterogeneity test

```{r eval = TRUE, echo = FALSE}
I2_all_data<-i2_ml(m)
I2_all_data

round(I2_all_data, digits = 3)

m_het <- rma.mv(yi = yi, 
                      V = vi, 
                      mods = ~ 1, 
                      random = list(~1|ID.article,
                                    ~1|ID.observation,
                                    ~1|Self.reported.or.measured,
                                    ~1|Specific.productivity.measure,
                                    ~1|Broad.productivity.measure),
                      data=all_data)
                
I2_m_het<-i2_ml(m_het)
round(I2_m_het, digits = 1) 


```

## Publication bias

### Figure 6
Funnel plots.

```{r eval = TRUE, echo = TRUE, fig.width=7, fig.height=7}
png("figure6.png", width = 205, height = 205, units='mm', res = 300)
  plot(1:10, 1:10, mar=c(3.5,0,3,0))
par(mfrow = c(2, 1))

selfreported_data<-subset(all_data, all_data$Self.reported.or.measured == "Self-reported")
meta_selfreported <- rma.mv(yi=yi, V=vi, data=as.data.frame(selfreported_data))
f1<-funnel(meta_selfreported, level=c(90, 95, 99), shade=c("white", "gray", "darkgray"),
       yaxis="seinv",xlab="Effect size",ylab="Precision (1/SE)",digits=c(1,0),xlim=c(-1.5,1.5),ylim=c(1,180), legend=TRUE,back="white", hlines = "white", main="a) Survey responses")

f1

meta_measured <- rma.mv(yi=yi, V=vi, data=as.data.frame(measured_data))
f2<-funnel(meta_measured, level=c(90, 95, 99), shade=c("white", "gray", "darkgray"),
       yaxis="seinv",xlab="Effect size",ylab="Precision (1/SE)",digits=c(1,0),xlim=c(-1.5,1.5),ylim=c(1,180), legend=TRUE,back="white", hlines = "white", main="b) Article output") 
f2

dev.off()


```

Multilevel meta-regression test for publication bias. Significant positive slope would suggest small-study effects (small-studies with larger effect sizes being published that skew my meta-analysis) . 

```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}
# Application of Equation 24 from the main manuscript of Nakagawa et. al 2021


publication.bias.model.r.all.se <- rma.mv(yi, vi,
                                          mods=~Variance.as.standard.error+Self.reported.or.measured -1,   
                                          random=list(~1|ID.article/ID.observation),
                                          data=all_data)


summary(publication.bias.model.r.all.se)

```
##Sensitivity analyses
Repeat analysis excluding the studies that use the following less direct measured of productivity: Research time (N = 4), job-loss (N = 1), burnout (N = 1), and number of projects (N = 1). It looks like excluding these does not change the results much.
```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}
# Create dataframe without less direct measures of productivity
sensitive_data<-subset(all_data, all_data$Broad.specific.productivity!="Job loss")
sensitive_data<-subset(sensitive_data, Broad.specific.productivity!="Burn-out")
sensitive_data<-subset(sensitive_data, Broad.specific.productivity!="Work hours")
sensitive_data<-subset(sensitive_data, Broad.specific.productivity!="Projects")
sensitive_data<-subset(sensitive_data, Broad.specific.productivity!="Job loss")
```
### Repeat model 1 (all effect sizes)
```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}
sm1<-rma.mv(yi, vi, random=~1|ID.article/ID.observation, data=sensitive_data)
sm1
summary(sm1)
```


### Repeat model 2 (separated by study type)
```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}

sensitive_data$Broad.productivity.measure[sensitive_data$Broad.productivity.measure=="Other"]<-"Survey"

tapply(sensitive_data$ID.observation, sensitive_data$Broad.productivity.measure, length)

sensitive_data$Broad.productivity.measure.reordered<-factor(sensitive_data$Broad.productivity.measure, c("Submissions", "Publications", "Survey"))

sm.area <- rma.mv(yi, vi, mods= ~ Broad.productivity.measure.reordered, random=~1|ID.article/ID.observation, data=sensitive_data)
sm.area
summary(sm.area)

sm.area1 <- rma.mv(yi, vi, mods= ~ Broad.productivity.measure.reordered-1, random=~1|ID.article/ID.observation, data=sensitive_data)
sm.area1

sp1 <- my.orchard(sm.area1, mod="Broad.productivity.measure.reordered", group = "Article.ID", data=sensitive_data, xlab = "Effect size", whisker=0.05,transfm = "none")+
     annotate("text", size=6, y=0.62, x=3.15, label=paste("-0.193 [-0.273, -0.113]"))+
     annotate("text", size=6, y=0.62, x=2.15, label=paste("-0.046 [-0.082, -0.009]"))+
     annotate("text", size=6, y=0.62, x=1.15, label=paste("-0.039 [-0.076, -0.001]"))+
  scale_fill_manual(values=c("#785EF0","#DC267F","#FE6100"))

sp1 

ggsave("sfigure2.png", width=10, height=10, dpi=300)
```
### Leave-one-out model
```{r eval = TRUE, echo = TRUE, fig.width=9, fig.height=9}
m<-rma.uni(yi, vi, random=~1|ID.article/ID.observation, data=all_data)
m
l1o<-leave1out(m)
forest(l1o$estimate, sei=l1o$se, header=TRUE, xlab="Leave One Out Estimate", refline=coef(res),xlim=c(-39,24))
```

